---
title: "Linear_Correlations"
output: html_document
---
---
title: "Linear_Correlations"
output: html_document
---

```{r,  include=FALSE}
library(ggplot2)    
library(tidyverse)  
library(gridExtra)
library(mlbench)
library(tidyverse)
library(ggmap)
library(gplots)
library(GGally)
library(lubridate)
knitr::opts_chunk$set(tidy = FALSE, message=FALSE, warning=FALSE)
options(htmltools.dir.version = FALSE)
library(rgdal)
library(maptools)

library(leaps)
library(caret)
library(lubridate)
library(mosaicData)
library(glmnet)
```

```{r}
daily <- read.csv('../../Data/1999_2021/aggragate_daily_data_2000to2020.csv')
names(daily)
```

```{r}
lowerFn <- function(data, mapping, method = "lm", ...) {
  p <- ggplot(data = data, mapping = mapping) +
    geom_point(colour = "blue") +
    geom_smooth(method = method, color = "red", ...)
  p
}
```

**Creating correlation matrix with ggpairs**
```{r}
pairs_plot <- ggpairs(
  daily[,3:11], lower = list(continuous = wrap(lowerFn, method = "lm")),
  diag = list(continuous = wrap("barDiag", color = "blue")),
  upper = list(continuous = wrap("cor", size = 2))
)
pairs_plot
```

**Creating table of correlations and removing redundancies**

```{r}
daily_trunc <- daily %>%
  select(-X, -date)

corr_table <- cor(daily_trunc) %>%
  as_tibble(rownames = "var1") %>%
    pivot_longer(cols = -var1, names_to = "var2", values_to = "coeff") %>%
    filter(var1 != var2) %>%
    arrange(desc(abs(coeff))) 

corr_table <- distinct(corr_table, coeff, .keep_all = TRUE)
corr_table
```

**Correlations of all variables to energy generated**

```{r}
energy_gen_corr <- cor(daily_trunc) %>%
  as_tibble(rownames = "var1") %>%
    pivot_longer(cols = -var1, names_to = "var2", values_to = "coeff") %>%
    filter(var1 == "energy_generated") %>%
    arrange(desc(abs(coeff))) 
energy_gen_corr
```

**Shifting influent flow, water level, effluent flow, turbine flow back and finding correlations with energy generation**

```{r}
lagged_daily <- daily %>%
  select(date, influent_flow, water_level, effluent_flow, energy_generated) %>%
  mutate(influent_flow_1 = lag(influent_flow, 1, order_by = date),
         influent_flow_3 = lag(influent_flow, 3, order_by = date),
         influent_flow_7 = lag(influent_flow, 7, order_by = date),
         influent_flow_21 = lag(influent_flow, 21, order_by = date),
         water_level_1 = lag( water_level, 1, order_by = date),
         water_level_3 = lag( water_level, 3, order_by = date),
         water_level_7 = lag( water_level, 7, order_by = date),
         water_level_21 = lag(water_level, 21, order_by = date),
         effluent_flow_1 = lag(effluent_flow, 1, order_by = date),
         effluent_flow_3 = lag(effluent_flow, 3, order_by = date),
         effluent_flow_7 = lag(effluent_flow, 7, order_by = date),
         effluent_flow_21 = lag(effluent_flow, 21, order_by = date)) %>%
  select(-date)


lagged_energy_gen_corr <- cor(drop_na(lagged_daily)) %>%
  as_tibble(rownames = "var1") %>%
    pivot_longer(cols = -var1, names_to = "var2", values_to = "coeff") %>%
    filter(var1 == "energy_generated") %>%
    arrange(var2) 
lagged_energy_gen_corr

```

# LASSO on additive model
```{r}
lasso_model <- train(
  energy_generated ~ ., data = daily_trunc,
  method = "glmnet",
  trControl = trainControl(method = "cv", number = 10),
  tuneGrid = data.frame(alpha = 1, lambda = 10^seq(-5, 1, length = 1000)),
  metric = "MAE"
)
```

```{r}
# Plot coefficients for each LASSO
plot(lasso_model$finalModel, xvar = "lambda", label = TRUE, col = rainbow(11))
```
```{r}
plot(lasso_model)
coef(lasso_model$finalModel, lasso_model$bestTune$lambda)
```

```{r}
# lasso_model$results gives metrics for every LASSO at every lambda
#lasso_model$results

# Just focus on those that correspond to the best lambda
lasso_model$results %>% 
  filter(lambda == lasso_model$bestTune$lambda)
```

Note: suspiciously high Rsquared. Possibly because turbine flow is directly related to energy generated. Attempt LASSO without these variables? Attempt LASSO with lagged variables?

**Running LASSO without poured flow, turbine flow, effluent flow, or water level**

```{r}
lasso_model_1 <- train(
  energy_generated ~ energy_charge + influent_flow + volume_used + energy_stored + maximum_demand, data = daily_trunc,
  method = "glmnet",
  trControl = trainControl(method = "cv", number = 10),
  tuneGrid = data.frame(alpha = 1, lambda = 10^seq(-5, 1, length = 1000)),
  metric = "MAE"
)

# Plot coefficients for each LASSO
plot(lasso_model_1$finalModel, xvar = "lambda", label = TRUE, col = rainbow(5))

plot(lasso_model_1)

coef(lasso_model_1$finalModel, lasso_model_1$bestTune$lambda)

lasso_model_1$results %>% 
  filter(lambda == lasso_model_1$bestTune$lambda)
```

**Notes**: Moderately significant Rsquared after removing obvious variables.

**Running LASSO with 7 day lag**

```{r}
daily_lagged_7 <- daily %>%
  mutate(energy_charge_7 = lag(energy_charge, 7, order_by = date),
         effluent_flow_7 = lag(effluent_flow, 7, order_by = date),
         influent_flow_7 = lag(influent_flow, 7, order_by = date),
         volume_used_7 = lag(volume_used, 7, order_by = date),
         poured_flow_7 = lag(poured_flow, 7, order_by = date),
         water_level_7 = lag(water_level, 7, order_by = date),
         energy_stored_7 = lag(energy_stored, 7, order_by = date),
         maximum_demand_7 = lag(maximum_demand, 7, order_by = date),
         ) %>%
  select(-date, -energy_charge,-effluent_flow, -influent_flow, -volume_used, -energy_stored, -poured_flow, -water_level, -energy_stored, -maximum_demand) %>%
  drop_na()

lasso_model_7 <- train(
  energy_generated ~ ., data = daily_lagged_7,
  method = "glmnet",
  trControl = trainControl(method = "cv", number = 10),
  tuneGrid = data.frame(alpha = 1, lambda = 10^seq(-5, 1, length = 1000)),
  metric = "MAE"
)

# Plot coefficients for each LASSO
plot(lasso_model_7$finalModel, xvar = "lambda", label = TRUE, col = rainbow(5))

plot(lasso_model_7)

coef(lasso_model_7$finalModel, lasso_model_7$bestTune$lambda)

lasso_model_7$results %>% 
  filter(lambda == lasso_model_7$bestTune$lambda)
```

*Note*: Including all variables at a 7 day lag results in a higher Rsquared than using a few of them with 0 lag, but LASSO doesn't eliminate any of them. Overfitted?

**Running LASSO without poured flow, turbine flow, effluent flow, or water level and lagging by 21 days**

```{r}
daily_lagged_21 <- daily %>%
  mutate(energy_charge_21 = lag(energy_charge, 21, order_by = date),
         effluent_flow_21 = lag(effluent_flow, 21, order_by = date),
         influent_flow_21 = lag(influent_flow, 21, order_by = date),
         volume_used_21 = lag(volume_used, 21, order_by = date),
         poured_flow_21 = lag(poured_flow, 21, order_by = date),
         water_level_21 = lag(water_level, 21, order_by = date),
         energy_stored_21 = lag(energy_stored, 21, order_by = date),
         maximum_demand_21 = lag(maximum_demand, 21, order_by = date),
         ) %>%
  select(-date, -energy_charge,-effluent_flow, -influent_flow, -volume_used, -energy_stored, -poured_flow, -water_level, -energy_stored, -maximum_demand) %>%
  drop_na()

lasso_model_21 <- train(
  energy_generated ~ ., data = daily_lagged_21,
  method = "glmnet",
  trControl = trainControl(method = "cv", number = 10),
  tuneGrid = data.frame(alpha = 1, lambda = 10^seq(-5, 1, length = 1000)),
  metric = "MAE"
)

# Plot coefficients for each LASSO
plot(lasso_model_21$finalModel, xvar = "lambda", label = TRUE, col = rainbow(5))

plot(lasso_model_21)

coef(lasso_model_21$finalModel, lasso_model_21$bestTune$lambda)

lasso_model_21$results %>% 
  filter(lambda == lasso_model_21$bestTune$lambda)
```

*Note*: significant drop in Rsquared, terrible estimate. 

# Conclusion

LASSO is not useful for predicting. Not surprising given the paper said that linear models are insufficient for optimizing hydroelectric dams. Probably wouldn't be useful for forecasting either.













